Containers and R: the Rockerverse and beyond

### 2019-08-12 manuscript.Rmd
Lori's rough draft paragraph style
  - Overview of Bioconductor
  - Bioconductor release version closely correlated with R release version
  - Bioconductor images build on top of rocker images
  - Tag Bioconductor images with R_ver_Bioc_ver for current/past accessibility
  - Built and distribute through DockerHub
  - Community contributed images

---
title: "Containers and R: the Rockerverse and beyond"
author:
  - name: Daniel Nüst
    affiliation: Institute for Geoinformatics, University of Münster 
    address:
    - Heisenbergstr. 2, 48149 Münster, Germany
    email: daniel.nuest@uni-muenster.de
  - name: Carl Boettiger
    #affiliation: -
    #email: -
  - name: Robrecht Cannoodt
    #affiliatin: -
    #email: -
  - name: Dirk Eddelbuettel
    #affiliation: -
    #email: -
  - name: Mark Edmondson
    #affiliation: -
    #email: -
  - name: Colin Fay
    #affiliation: -
    #email: -
  - name: Ben Marwick
    #affiliation: -
    #email: -
  - name: Karthik Ram
    #affiliation: -
    #email: -
  - name: Noam Ross
    #affiliation: -
    #email: -
  - name: Nan Xiao
    #affiliation: -
    #email: - 
abstract: >
  The Rocker project provides widely-used Docker images for R across different application scenarios.
  This articles surveys describe downstream projects building upon Rocker.
  We also look beyond Rocker to other projects connecting containerisation with R.
  These use cases and the diversity of applications demonstrate the power of Rocker and containerisation for collaboration, effectivity, scalability, and transparency.
preamble: >
   \usepackage{longtable}
output: rticles::rjournal_article
---

@@

@@

>>>>

---
title: "Containers and R: the Rockerverse and beyond"
author:
  - name: Daniel Nüst
    affiliation: Institute for Geoinformatics, University of Münster 
    address:
    - Heisenbergstr. 2, 48149 Münster, Germany
    email: daniel.nuest@uni-muenster.de
  - name: Carl Boettiger
    #affiliation: -
    #email: -
  - name: Robrecht Cannoodt
    #affiliatin: -
    #email: -
  - name: Dirk Eddelbuettel
    #affiliation: -
    #email: -
  - name: Mark Edmondson
    #affiliation: -
    #email: -
  - name: Colin Fay
    #affiliation: -
    #email: -
  - name: Ben Marwick
    #affiliation: -
    #email: -
  - name: Karthik Ram
    #affiliation: -
    #email: -
  - name: Noam Ross
    #affiliation: -
    #email: -
  - name: Nan Xiao
    #affiliation: -
    #email: - 
  - name: Lori Shepherd
    affiliation: Roswell Park Cancer Institute
    address:
    - Elm & Carlton Streets, Buffalo, New York, 14263 USA
    email: lori.shepherd@roswellpark.org
abstract: >
  The Rocker project provides widely-used Docker images for R across different application scenarios.
  This articles surveys describe downstream projects building upon Rocker.
  We also look beyond Rocker to other projects connecting containerisation with R.
  These use cases and the diversity of applications demonstrate the power of Rocker and containerisation for collaboration, effectivity, scalability, and transparency.
preamble: >
   \usepackage{longtable}
output: rticles::rjournal_article
---

@@

#
# Draft bullet point notes (brainstorming session)
#

@@

#
# Rough draft - they asked one sentance per line
#

Bioconductor is a free, open source and open development project aimed at the analysis and visualization of genetic and genomic data utilizing the R programming language [1]. 
The project currently consists of approximately 1741 R packages and has a semiannual release subsequent to the Comprehensive R Archive Networks release of R [2]. 
Biocondctor release versions are tightly correlated with a version of R. 
Docker images allow availability of current and past versions of R/Bioconductor for convenience and reproducibility.

The Bioconductor docker images are built on top of rocker/r-ver and rocker/rstudio. By utilizing rocker as a base image, the R version and majority of package system dependencies are already included in a Bioconductor base image. 
Bioconductor will install packages based on the selected R version and therefore leverage the rocker/rstudio and rocker/r-ver use of R version tagging. 
Bioconductor selects the desired version of R from rocker, installs BiocManager which will default to the appropriate Bioconductor version, and creates a Bioconductor docker image with a informative tag (R_version_Bioc_version). 
The images are built and distributed to the community through DockerHub [3]. 
All past and current combinations of R and Bioconductor are therefore accessible via a specific docker tag.

The Bioconductor base image contains the minimial system depenedencies and R from rocker, and the BiocManager package for installation of packages configured for the proper version of Bioconductor given the installed R version. 
Bioconductor has several other docker images specific for various areas of research. 
The Bioconductor core image installs the most downloaded Bioconductor packages based on collected statistics. 
The other Bioconductor images are community maintained: Laurent Gatto maintains proteomics images, Steffen Neumann maintains metabolomics images, and the RGLab maintains cytometry images. 
All community maintained images build on top of the Bioconductor base image and therefore indirectly the rocker images [4].

#
# Need to move these to bib file and properly format. 
#

[1]  Gentleman, Robert C.; Carey, Vincent J.; Bates, Douglas M.; Bolstad, Ben;
Dettling, Marcel; Dudoit, Sandrine; Ellis, Byron; Gautier, Laurent; Ge,
Yongchao; Gentry, Jeff; Hornik, Kurt; Hothorn, Torsten; Huber, Wolfgang; Iacus,
Stefano; Irizarry, Rafael; Leisch, Friedrich; Li, Cheng; Maechler, Martin;
Rossini, Anthony J.; Sawitzki, Gunther; Smith, Colin; Smyth, Gordon; Tierney,
Luke; Yang, Jean Y. H.; Zhang, Jianhua (2004). "Bioconductor: open software
development for computational biology and bioinformatics". Genome Biology. 5
(10): R80. doi:10.1186/gb-2004-5-10-r80. PMC 545600

[2] R Core Team (2019). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria.
  URL https://www.R-project.org/

[3] ??DockerHub citation?

[4] ??Reference our docker help page/github?
